---
title: "Augury — Lead-Lag Analysis"
subtitle: "Does social stance carry information about prediction-market prices?"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    theme: cosmo
execute:
  warning: false
  message: false
---

```{r setup}
#| include: false
for (file in list.files(file.path("..", "R"), pattern = "\\.R$", full.names = TRUE)) {
  source(file)
}
library(dplyr)
library(ggplot2)
library(knitr)

MODEL_VERSION <- Sys.getenv("AUGURY_MODEL_VERSION", "vader-3.3.2")
ALPHA <- 0.05
RUN_ID <- format(Sys.time(), "%Y%m%dT%H%M%SZ")
```

## What this report claims, and what it does not

This is a **calibration and causality audit**, not a trading recommendation. It
asks one question — does the social signal `S(t)` tell us anything about where a
prediction-market price goes next that the price's own history does not already
tell us — and reports the answer including when that answer is "no".

Three guards are applied to everything below, because without them the analysis
would reliably manufacture findings:

- **Logit transform.** Prices are bounded in $[0,1]$. A market at 0.03 cannot
  fall much further, so its innovations are asymmetric and its level is
  near-certainly non-stationary in the sense the Granger test assumes.
- **Stationarity is tested, not assumed.** An ADF result travels with every
  test. Granger causality between two integrated series produces impressive
  p-values from spurious regression alone.
- **Multiple comparisons.** Testing dozens of markets at $p < 0.05$ yields
  roughly one "significant" result per twenty from chance. Every p-value here is
  Benjamini-Hochberg corrected **across the whole batch**, and significance is
  defined off the adjusted value only.

```{r connect}
#| include: false
db_available <- augury_available()
```

```{r no-db}
#| eval: !expr "!db_available"
#| results: asis
cat("::: {.callout-warning}\n")
cat("## Database unavailable\n\n")
cat("Could not connect to TimescaleDB, so this report has no data to analyze.\n\n")
cat("Start the stack and populate it:\n\n")
cat("```\nmake up && make db-migrate\naugury sync-config && augury refresh\n```\n")
cat(":::\n")
```

```{r load}
#| eval: !expr db_available
#| include: false
connection <- augury_connect()
markets <- fetch_markets(connection)

market_data <- list()
market_meta <- list()

for (i in seq_len(nrow(markets))) {
  market_id <- markets$market_id[i]

  signals <- fetch_signals(connection, market_id, MODEL_VERSION)
  prices <- fetch_prices(connection, market_id)

  if (nrow(signals) == 0 || nrow(prices) == 0) next

  aligned <- align_on_grid(
    signals %>% select(ts, s_t),
    prices %>% select(ts, mid),
    price_col = "mid"
  )

  if (nrow(aligned) < 20) next

  market_data[[market_id]] <- aligned
  market_meta[[market_id]] <- markets[i, ]
}
```

```{r analyze}
#| eval: !expr db_available
#| include: false
results <- if (length(market_data) > 0) {
  analyze_markets(market_data, alpha = ALPHA)
} else {
  list()
}
results_frame <- results_to_frame(results)
```

## Coverage

```{r coverage}
#| eval: !expr db_available
tibble::tibble(
  Metric = c(
    "Tracked markets",
    "Markets with enough aligned history",
    "Model version",
    "Minimum observations required"
  ),
  Value = c(
    nrow(markets),
    length(market_data),
    MODEL_VERSION,
    "20 aligned hourly bars"
  )
) %>% kable()
```

```{r insufficient}
#| eval: !expr "db_available && length(market_data) == 0"
#| results: asis
cat("::: {.callout-important}\n")
cat("## Not enough history yet\n\n")
cat("No market has 20 aligned hourly observations of both signal and price.\n")
cat("Granger causality on less than that is not interpretable, so no test was run.\n\n")
cat("Let `augury-ingest` accumulate data for a few days and re-render.\n")
cat(":::\n")
```

## Granger causality: does the signal lead price?

```{r granger-table}
#| eval: !expr "db_available && length(market_data) > 0"
results_frame %>%
  select(
    Market = market_id,
    Lag = lag_order,
    N = n_obs,
    F = f_statistic,
    `p (raw)` = p_value,
    `p (FDR)` = p_value_adj,
    Significant = significant,
    `ADF signal` = adf_p_signal,
    `ADF price` = adf_p_price
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
  arrange(`p (FDR)`) %>%
  kable(caption = paste0(
    "Signal -> price. Significance is FDR-corrected at alpha = ", ALPHA,
    ". ADF p-values below 0.05 indicate a stationary (testable) series."
  ))
```

```{r verdict}
#| eval: !expr "db_available && length(market_data) > 0"
#| results: asis
n_significant <- sum(results_frame$significant, na.rm = TRUE)
n_tested <- nrow(results_frame)
n_nonstationary <- sum(
  results_frame$adf_p_signal >= 0.05 | results_frame$adf_p_price >= 0.05,
  na.rm = TRUE
)

if (n_significant == 0) {
  cat("::: {.callout-note}\n")
  cat("## Verdict: no evidence that the signal leads price\n\n")
  cat(sprintf(
    "Across %d market(s), none survived FDR correction at alpha = %.2f.\n\n",
    n_tested, ALPHA
  ))
  cat("This is a legitimate and informative result. It is consistent with the ")
  cat("markets being efficient with respect to public social sentiment — which ")
  cat("is what one should expect from a liquid market, and is precisely the ")
  cat("hypothesis this project was built to test rather than to confirm.\n")
  cat(":::\n")
} else {
  cat("::: {.callout-important}\n")
  cat(sprintf("## %d of %d market(s) show a significant lead\n\n", n_significant, n_tested))
  cat("Before treating this as real, check each of the following:\n\n")
  cat("- Does the ADF column show both series were stationary? If not, the ")
  cat("result may be spurious regression.\n")
  cat("- Is `N` large enough that the lag order was identifiable?\n")
  cat("- Is the effect present on the market's *quiet* periods, or only around ")
  cat("one news event that moved sentiment and price simultaneously?\n")
  cat("- Would it survive out of sample, on markets not used to choose the ")
  cat("specification?\n")
  cat(":::\n")
}

if (n_nonstationary > 0) {
  cat("\n::: {.callout-warning}\n")
  cat(sprintf(
    "%d market(s) had at least one non-stationary series (ADF p >= 0.05). Treat their p-values with suspicion.\n",
    n_nonstationary
  ))
  cat(":::\n")
}
```

## Reverse direction: does price lead the signal?

A signal that merely *follows* price is the null result worth being able to
report. If the reverse direction is stronger than the forward one, the pipeline
is measuring people reacting to prices they have already seen.

```{r reverse}
#| eval: !expr "db_available && length(market_data) > 0"
reverse_results <- list()
for (market_id in names(market_data)) {
  data <- market_data[[market_id]]
  outcome <- tryCatch(
    {
      test <- granger_test(data$signal, data$price, direction = "price_to_signal")
      test$market_id <- market_id
      test
    },
    error = function(e) NULL
  )
  if (!is.null(outcome)) reverse_results[[market_id]] <- outcome
}

reverse_frame <- results_to_frame(apply_fdr_correction(reverse_results, alpha = ALPHA))

if (nrow(reverse_frame) > 0) {
  reverse_frame %>%
    select(Market = market_id, Lag = lag_order, N = n_obs,
           F = f_statistic, `p (FDR)` = p_value_adj, Significant = significant) %>%
    mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
    arrange(`p (FDR)`) %>%
    kable(caption = "Price -> signal, FDR-corrected.")
}
```

## Per-market detail

```{r per-market}
#| eval: !expr "db_available && length(market_data) > 0"
#| results: asis
#| fig-width: 9
#| fig-height: 4
for (market_id in names(market_data)) {
  aligned <- market_data[[market_id]]
  meta <- market_meta[[market_id]]

  cat("\n\n### ", market_id, "\n\n")
  cat(meta$title, "\n\n")

  print(plot_signal_vs_price(aligned, title = market_id))

  ccf_table <- cross_correlation(aligned$signal, aligned$price, max_lag = 12)
  print(plot_cross_correlation(ccf_table))
}
```

## Calibration

```{r calibration}
#| eval: !expr db_available
#| results: asis
scores <- fetch_scores(connection)

if (nrow(scores) == 0) {
  cat("::: {.callout-note}\n")
  cat("No markets have resolved yet, so no Brier scores exist.\n\n")
  cat("A Brier Skill Score can only be computed against a realized outcome; ")
  cat("until a tracked market settles, calibration is unmeasurable.\n")
  cat(":::\n")
} else {
  print(
    scores %>%
      select(Market = market_id, Model = model_version, N = n_obs,
             Calibration = calibration, `BS signal` = brier_signal,
             `BS market` = brier_market, BSS = bss, Outcome = resolved_outcome) %>%
      mutate(across(where(is.numeric), ~ round(.x, 4))) %>%
      kable(caption = "BSS > 0 means the signal beat the market's own price as a forecast.")
  )
}
```

```{r write-back}
#| eval: !expr "db_available && length(market_data) > 0"
#| include: false
write_granger_results(connection, results_frame, RUN_ID, MODEL_VERSION)
```

```{r disconnect}
#| eval: !expr db_available
#| include: false
DBI::dbDisconnect(connection)
```

## Reading this correctly

A significant Granger result means the signal's history improves a linear
forecast of price beyond price's own history, on this sample, under this
specification. It does **not** establish a causal mechanism, and it is not
tradable on its own: it says nothing about transaction costs, about the spread
that would be crossed to act on it, or about whether the relationship persists
once anyone does.

The more likely outcome for a liquid market is no relationship at all. That
result is worth having, and this report is built to state it plainly rather than
to search for a specification that avoids it.
